---
title: "Ghostrunner"
subtitle: "Original Soundtrack"
author: "Rich Leyshon."
date: last-modified
description: "A review of the soundtrack to the 2020 hit video game
Ghostrunner, composed by Daniel Deluxe."
categories:        
    - Electronic
    - Synthwave
    - Video Game Music
image: https://i.scdn.co/image/ab67616d0000b2736bbb199a6537fe7d858ec523
image-alt: Album cover art
toc: true
---

## Introduction

I'm thrilled to kick off the music review section of my blog with a highly
engaging work by one of my favourite synth artists.

::: {.callout-note}
This is a record with high turnover & great energy - one to put on loop when
you need to pull out all the stops on a thorny problem.
:::

Ghostrunner was one of the critically acclaimed independent platformer games of
2020. It's aged particularly well on the latest generation of consoles, thanks
in no small part to its fantastic soundtrack. The developers made a safe bet
in selecting Daniel Deluxe to compose the nihilistic, cyberpunk soundtrack.
Deluxe has been producing knockout darksynth since 2014 and has been a regular
feature of my working music since I started programming. 

Deluxe's 2014 effort by the name of
[Darkness](https://open.spotify.com/album/1G66EjFlNMcXL4sudDiYEF) was one of
the first synthwave songs I had come across and partly the reason why I have
gone so deep with this genre. It's a fantastic piece of 80s romantic nostalgia
with evocative speech samples taken from the 1985 fantasy movie Legend,
starring Tim Curry as the eponymous Darkness - a character that haunted the
nightmares of many '80s children. Much of Deluxe's catalogue is worthy of your
time and Ghostrunner OST is possibly the crowning achievement. As his fifth and
latest (at time of writing) studio album, Ghostrunner draws and builds upon the
stylistic themes explored in previous albums. Though in this effort, a shorter
average track length has kept things laser-focussed. Ghostrunner is a great
introduction to synthwave of the catalogue of Daniel Deluxe as a whole, but
luckily for the newcomer - there's plenty of depth to mine from this musical
vein.

## Music for Cyborg Ninjas

It's worth noting that this music isn't for everyone. But I've never favoured
that sort of music myself and wouldn't find the time to review it. This music
was written to energise players ready to engage in parkour-slash-em-up fun, 
without distracting from the brief interludes of narrative. As with
all of my favoured music for work, you won't find any ballads or warbling
vocals here, so if you wan't something to humm along to, you'd best find
another blog!

But if you're down for this, I'd recommend clicking play on the spotify player
button and enjoying the tunes while you read.

<iframe style="border-radius:12px" src="https://open.spotify.com/embed/album/6mS0ssCxtLvB0IcVP7sR66?utm_source=generator" width="40%" height="80" frameBorder="0" allowfullscreen="" allow="autoplay; clipboard-write; encrypted-media; fullscreen; picture-in-picture" class=center></iframe>
<br>
Let's take a look at the audio feature analysis for this album. For a formal
description of the audio feature categories presented here, please consult the 
<a href="https://developer.spotify.com/documentation/web-api/reference/get-audio-features" target="_blank">Spotify Developer API Documentation</a>.

```{r, width="300px", fig.align="center"}
#| warning: false
#| echo: false
library(here)
library(kableExtra)
library(glue)
source(here("R/music-reviews/spotify-album.R"))
source(here("R/music-reviews/config-secrets.R"))

viz_album_audio_features(
    title_font_size = 16,
    subtitle_font_size = 12,
    title_hjust = 1.8,
    subtitle_hjust = 0.1,
    debug = FALSE
)

```

::: {.callout-tip collapse="true"}
## Interpreting the Ridge Plot (Click to expand)

These ridge plots show a histogram of the mean audio feature values across each track in the album. Look across the horizontal axis at where the peaks for each feature occur. For eample, if the album contains lots of tracks with busy vocals, the speechiness feature will show a peak to the right of the horizontal axis. Conversely, an instrumental album would have a speechiness peak to the left of the horizontal axis but you would also expect it to have a peak to the right of the axis in the instrumentalness feature.
:::

Ghostrunner OST is clearly an instrumental album. This plot show that the tracks all score lowly on speechiness and highly on instrumentalness, with tight distributions. The album generally displays high measures for energy and danceability, though with broader distributions than that displayed by speechiness & instrumentalness. The majority of the album is characterised by clear, steady and fast rhythm. Interestingly, the acousticness feature presents a very broad range of distribution in the confidence that tracks do not have electrical amplification. This really does not seem correct to me, as the album is unapologetically synth all the way. This has potentially been affected by production effects favouring reverb and echo, but that is surmising on my part. Lastly, I'll turn to valence, which is a measure of positivity within the music. This feature is left-skewed. A higher number of lower valence values makes sense, as much of the album carries a sinister, Orwellian, digital-dystopian tone - generally interpretable as negative I would agree. The presence of higher values in this feature suggests that a fewer number of the tracks may be characterised with positive emotion. 

## The Tracks

Below are the mean audio feature values for each track. The last row (green)
presents a mean summary of the album.

```{r}
#| echo: false
tab = create_album_audio_summary_table()
# create album average
numeric_cols = select(tab, where(is.numeric))
summary_tab = summarise(numeric_cols, across(everything(), mean))
summ_row = cbind(track_name = glue::glue("Album Mean"), summary_tab)
tab = rbind(tab, summ_row)
# Visualise
summ_ind = nrow(tab)
kableExtra::kbl(tab) %>%
row_spec(summ_ind, color = "white", background = "#146542")


```

**Air**
Beautiful synth bleeparrangement overlayed with menacing robot groans. A mashy
bass beat rounds out this atmospheric ode to a mysterious, futuristic
cityscape.


**Infiltrator**
Bassy experiments with tempo against a stripped-back rhythm section. Reminds me
of scrap brain zone on sonic the hedgehog. Intermittent alert signals and
robotic glitches. Glimpses of some wonderful serene motif before being thrust
back into the main formation.

**Blood and steel**
This track is very recognisable. Inspires an unsettling feeling of De ja vú
within the listener. Conjuring images of a futuristic crime scene
investigation. A dreamy intro prior to a strong drum & bass injection. The
track culminates in a growing phaser storm that smothers the rest of the track
before it cuts at the point of its culmination. A highlight of the album.

**Dharma**
Anyone of a certain age will undoubtedly be reminded of ABC's gripping scifi 
thriller Lost, of the early noughties. This track strikes a noticeably menacing
tone. A dripping, ticking lick loops over a bossy bass hook. Lazer phasers
growing in tone & pitch throughout for good measure. 

**Capture**
A zoned out, decidedly creepy intro with an android snigger looped over bell
tolls, metallic pipes clanging and a menagerie of industrial noise. A slower
beat chugging away only to take frequent breaks to peak at you from the
shadows. Sinister but with style.

**Forget the past**
Begins on a melancholy reflection that seems to consider the song's title some 
advice to self. Continues the chilling set by the previous track with an
inhumane battery of bass noise, giving way to serene, elongated spacetrips of
harmonic synth.

**Access denied**
A collective of error signals experienced through a thick gelatinous membrane.
This song feels like being rudely awoken from the matrix. An electric wire to
the brain in the high tempo intro of a squelchy phaser bleep on loop. Drum &
cymbal snares introduces some peril to the experience. Another favourite of
mine. 

**Let them know** - higher range melodic expletives over a stripped back drum beat, and a growing, bubbling, frothy arsenal of phaser barrages. Giving way to a much stronger bass and dipping in and out with further volleys. Wipeout style high speed wipe and some time out ticks, sounds like hitting a spoon on a tin bowl. Robot timeout. Layers wiped together in a busy, structured crescendo.

**Sector** - mysterious hyper speed contrail sound. With some product ionised robot intonation, rapping over the crunchy bass drum. Imagine you programmed a robot with the licence to low-fi rap. 

**Factory** - clever mix of major minor android aria signals. Locking in some bassy robo groans with some mysterious echo -flared laser trills. Descends into a mission impossible repeato added some beautiful high freq synth melody.

**Striker** - enters with an industrial explosion experienced from a vast distance, echoing out into a wide pool of synth intergration tests, like circuit fingers working through a mossy undertow. High freq synth chimes with glitch notation slips away to a seedy, menacing robo bridge. The pretty synth returns to work over the menacing robo.

**Celerity** - Prodigy-like industrial D&B with some Charlie always tells his mommy fairground loop. Being glitched out with a jarring, chugging interference. Giving way to some graceful tech wipes for a moment before the interference invades once more. Moving to a double time march to up the ante, long distance robo chase,.

**The orb** - shimmering, robot talk phasing in and out of consciousness. Moves onto experiment with the beat and tone, swaying in and out of the fore as the beat establishes command. 

**Truth to power** - a whispered android complaint grows and takes full form, before being resigned to radio playback. Jumps in and out of the foreground while comets wipe by and leaves a trail of icy debris zooming past the listener. 

**Blaster** - something larger rising out of the depths, waving away in front of your ears. Once surfaced, reveals itself to be largely drum-driven automation with a bassy tempest of brass billowing away under the surface. 

**Razor** - accusatory synth calls against a halloween-esque synth sustain. Emerges onto an 80s horror set before adopting a atention-demanding rhythm section. High-end synth parrots away while some os-level tech blurb ramps up to the fore. 

**Vendetta** - reliable, robot trill dances between pitches, laying off for some harmonious electro organ to swing in and out overhead. While machine gun volleys show their weight to the beat.

**Air** (Cybervoid) a welcome reprise to the first track revisits the suspicious first day in a cybernetic future. Blazing bass warmly intervenes the laser hook and meshes around as the track progresses. 

**Solitude** - dreamy space age synthetic  Bach - rises from the murky depths to sit in a shimmering orchestral pool while sum squishy, rubber deep bass robe vocals sound out over the morass. 

**Sundown** - gothic android monk prayer gives way to laser piano overlays. Cutting out to flatlined electro strings & the futuristic remnants of whatever birds evolved into once AI conquered the Earth. 

## Highs and Lows

## Overall