---
title: "Ghostrunner"
subtitle: "Original Soundtrack"
author: "Rich Leyshon"
date: September 5 2023
description: "A review of the soundtrack to the 2020 hit video game
Ghostrunner, composed by Daniel Deluxe."
categories:
- Electronic
- Synthwave
- Video Game Music
image: https://i.scdn.co/image/ab67616d0000b2736bbb199a6537fe7d858ec523
image-alt: Album cover art
toc: true
---

## Introduction

I'm thrilled to kick off the music review section of my blog with an engaging
work by one of my favourite synth artists.

::: {.callout-note}
This is a record with high turnover & great energy - one to put on loop when
you need to pull out all the stops on a thorny problem.
:::

Ghostrunner was one of the acclaimed independent platformer games of
2020. It's aged particularly well on the latest generation of consoles, thanks
in no small part to its fantastic soundtrack. The developers made a safe bet
in selecting Daniel Deluxe to compose the nihilistic, cyberpunk soundtrack.
Deluxe has been producing knockout darksynth since 2014 and has been a regular
feature of my working music since I started programming.

Deluxe's 2014 effort by the name of
[Darkness](https://open.spotify.com/album/1G66EjFlNMcXL4sudDiYEF) was one of
the first synthwave songs I had come across and partly the reason why I have
gone so deep with this genre. It's a fantastic piece of 80s romantic nostalgia
with evocative speech samples taken from the 1985 fantasy movie Legend,
starring Tim Curry as the eponymous Darkness - a character that haunted the
nightmares of many '80s children. Much of Deluxe's catalogue is worthy of your
time and Ghostrunner OST is the crowning achievement. As his fifth and
latest (at time of writing) studio album, Ghostrunner draws and builds upon the
stylistic themes explored in previous albums. Though in this effort, a shorter
average track length has kept things laser-focussed. Ghostrunner is a great
introduction to synthwave of the catalogue of Daniel Deluxe as a whole, but
luckily for the newcomer - there's plenty of depth to mine from this musical
vein.

## Music for Cyborg Ninjas

It's worth noting that this music isn't for everyone. But I've never favoured
that sort of music myself and wouldn't find the time to review it. This music
was written to energise players ready to engage in parkour-slash-em-up fun,
without distracting from the brief interludes of narrative. As with
all my favoured music for work, you won't find any ballads or warbling
vocals here, so if you want something to hum along to, you'd best find
another blog!

But if you're down for this, I'd recommend clicking play on the spotify player
button and enjoying the tunes while you read.

<iframe style="border-radius:12px" src="https://open.spotify.com/embed/album/6mS0ssCxtLvB0IcVP7sR66?utm_source=generator" width="40%" height="80" frameBorder="0" allowfullscreen="" allow="autoplay; clipboard-write; encrypted-media; fullscreen; picture-in-picture" class=center></iframe>
<br>
Let's take a look at the audio feature analysis for this album. For a formal
description of the audio feature categories presented here, please consult the
<a href="https://developer.spotify.com/documentation/web-api/reference/get-audio-features" target="_blank">Spotify Developer API Documentation</a>.

```{r, width="300px", fig.align="center"}
#| warning: false
#| echo: false
library(here)
library(kableExtra)
library(glue)
source(here("R/music-reviews/spotify-album.R"))
source(here("R/music-reviews/config-secrets.R"))

viz_album_audio_features(
title_font_size = 16,
subtitle_font_size = 12,
title_hjust = 1.8,
subtitle_hjust = 0.1,
debug = FALSE
)

```

::: {.callout-tip collapse="true"}
## Interpreting the Ridge Plot (Click to expand)

These ridge plots show a histogram of the mean audio feature values across each
track in the album. Look across the horizontal axis at where the peaks for each
feature occur. For example, if the album contains lots of tracks with busy
vocals, the speechiness feature will show a peak to the right of the horizontal
axis. Conversely, an instrumental album would have a speechiness peak to the
left of the horizontal axis but you would also expect it to have a peak to the
right of the axis in the instrumentalness feature.
:::

Ghostrunner OST is clearly an instrumental album. This plot shows that the
tracks all score lowly on speechiness and highly on instrumentalness, with
tight distributions. The album generally displays high measures for energy and
danceability, though with broader distributions than that displayed by
speechiness & instrumentalness. The majority of the album is clear, steady and
fast rhythm. Interestingly, the acousticness feature presents a very broad
range of distribution in the confidence that tracks do not have electrical
amplification. This really does not seem correct to me, as the album is
unapologetically synth all the way. This is potentially affected by production
effects favouring reverb and echo, but that is surmising on my part. Lastly,
I'll turn to valence, which is a measure of positivity within the music. This
feature is right-skewed. A higher number of lower valence values makes sense,
as much of the album carries a sinister, Orwellian, digital-dystopian tone -
generally interpretable as negative I would agree. The presence of higher
values in this feature suggests that a fewer number of the tracks may be
characterised with positive emotion.

## The Tracks

Below are the mean audio feature values for each track. The last row (green)
presents a mean summary of the album.

```{r}
#| echo: false
tab = create_album_audio_summary_table()
# create album average
numeric_cols = select(tab, where(is.numeric))
summary_tab = summarise(numeric_cols, across(everything(), mean))
summ_row = cbind(track_name = glue::glue("Album Mean"), summary_tab)
tab = rbind(tab, summ_row)
# Visualise
summ_ind = nrow(tab)
kableExtra::kbl(tab) %>%
row_spec(summ_ind, color = "white", background = "#146542")

```

**Air**
Beautiful synth bleep arrangement overlayed with menacing robot groans. A mashy
bass beat rounds out this atmospheric ode to a mysterious, futuristic
cityscape.

**Infiltrator**
Bassy experiments with tempo against a stripped-back rhythm section. Reminds me
of scrap brain zone on sonic the hedgehog. Intermittent alert signals and
robotic glitches. Glimpses of some wonderful serene motif before being thrust
back into the main formation. At the time of writing, this track is currently
the most popular on this album.

**Blood and steel**
This track is very recognizable. Inspires an unsettling feeling of Deja vú
within the listener. Conjuring images of a futuristic crime scene
investigation. A dreamy intro before a strong drum & bass injection. The
track culminates in a growing phaser storm that smothers the rest of the track
before it cuts at the point of its culmination. A highlight of the album.

**Dharma**
Anyone of a certain age will likely remember ABC's gripping sci fi
thriller Lost, of the early noughties. Dharma was the name of the malevolent. corporation behind much of the protagonists woes. This track strikes a fittingly menacing
tone. A dripping, ticking lick loops over a bossy bass hook. Lazer phasers
growing in tone & pitch throughout for good measure.

**Capture**
A zoned out, decidedly creepy intro with an android snigger looped over bell
tolls, metallic pipes clanging and a menagerie of industrial noise. A slower
beat chugging away only to take frequent breaks to peak at you from the
shadows. Sinister but with style.

**Forget the past**
Begins on a melancholy reflection that seems to consider the song's title some
advice to self. Continues the chilling set by the previous track with an
inhumane battery of bass noise, giving way to serene, elongated spacetrips of
harmonic synth.

**Access denied**
A collective of error signals experienced through a thick gelatinous membrane.
This song feels like being rudely awoken from the matrix. An electric wire to
the brain in the high tempo intro of a squelchy phaser bleep on loop. Drum &
cymbal snares introduce some peril to the experience. Another favourite of
mine.

**Let them know**
High pitch melodic trills over a stripped back drum beat with a growing,
bubbling, frothy arsenal of phaser barrages. The trills give way to a strong
bass line, dipping in and out of the foreground. Halfway in, a Wipeout-style
high speed fade is accompanied by checkpoint ticks that sound a bit like
hitting metal on metal. This track has a great time trial feel. The layers
wipe together in a busy, well-structured crescendo. This is the highlight of
the album for me.

**Sector**
Starting on a mysterious, hyper-speed contrail of sound. With some
productionised robot intonation rapping over the crunchy bass drum. Joined
later by a Japanese-inspired treble section for the interested Japanophile.
Reminiscent of the excellent
[Triads by Code Elektro](https://open.spotify.com/album/3K48lplYXbQSc87pfuJ0Pa)
(review sure to come at some point).

**Factory**
A clever mix of major and minor android aria signals. Locking in some bassy
robo groans with some mysterious echo-flared laser trills. The track descends
into a mission impossible repeato while adding some beautiful high frequency
synth melody.

**Striker**
Enters with an industrial explosion experienced from a vast distance, rippling
out into a wide pool of synth integration tests, like circuit fingers working
through a mossy undertow. High frequency synth chimes with a glitched notation
slips away to a seedy, menacing robotic bridge. The pretty synth
returns to work over the menacing robo. A close second favourite.

**Celerity** - Celerity is not a word I have encountered before, but according
to [dictionary.com](https://www.dictionary.com/browse/celerity) it means swift
movement. An interesting choice of title for this track as I interpret the
rhythm to be more steady.
Starting on a
[Prodigy-like industrial D&B](https://open.spotify.com/track/3xH7xicTlukrmnHIHINTW0)
replete with an accompanying fairground loop. A glitched-up affair with a
jarring, chugging interference playing with your expectations throughout. This
moves on to some graceful tech wipes for a moment before the interference
dominates once more. Ending on a commandeering double-time march to up the
ante, this track conjures a long-distance robotic pursuit.

**The orb**
At this point in the album, we encounter a run of shorter, more intense songs.
The orb has a brilliant, shimmering, robot talk phasing in and out of
consciousness. Moving on to experiment with beat and tone patterns, swaying in
and out of the fore as the beat establishes command. Short, but oh, so sweet.

**Truth to power**
A chunky drum section overlayed by a whispered android complaint growing
to take full form, before relegated to radio playback. The android voice
jumps in and out of focus while comets wipe by and leaves a trail of icy debris
zooming past the listener. Likely to be the sort of music that Chat-GPT
generates when frustrated by the lack of imagination presented by its users.

**Blaster**
Watch out for this one - there's something larger rising out of the depths,
waving away in front of your ears. Once surfaced, it reveals itself to be
largely drum-engined automation with a bassy tempest of brass billowing away
under the surface, like an automaton that's managed to preserve the very last
of the human tuba instrumentalists, lifespan extended with cybernetic implants.

**Razor**
Accusatory synth calls out against a halloween-esque synth sustain. Emerging
out onto an '80s horror set before adopting an attention-demanding rhythm.
High-end synth parrots away while some os-level tech blurb ramps up to the
fore.

**Vendetta**
A reliable, robotic trill dances between pitches, laying off for some
harmonious electro organ to swing in and out overhead. Machine gun volleys
shower their weight to the beat.

**Air (Cybervoid)**
A welcome reprise to the first track of the album revisits a suspicious first
day in a cybernetic future. A subtle difference to the original but more
immediate. Blazing and warm bass tones intervene the laser hook and encompasses
around the track as it progresses.

**Solitude**
A dreamy space-age synthetic Bach rises from the murky depths to sit in a
shimmering orchestral pool, composing his techno-symphonies. In stark contrast,
squishy, rubbery deep bass android vocals sound out over the alien morass.

**Sundown**
A gothic android monk prayer cedes to laser piano overlays. Cutting out to
flatlined electro strings & the futuristic remnants of whatever birds evolved
into once AI conquered the Earth. A contemplative end to a high-energy album.
Likely to reflect a sombre tone to the end of the video game, though I
cannot confirm this as I was pretty rubbish when I tried it.

## Highs and Lows

No real lows to talk of in this album, but the three standout tracks that I
recommend are:

1. Let them know
2. Striker
3. Blood and steel

## Overall

A smashing piece of dark synthpop that delivers focussed energy while
conserving your concentration. The theme is sinister at points but never loses
its potency for concentrainment. A good choice for smashing out some shallow
work or a briefing note in a short space of time. Perhaps not as inspiring or
uplifting as some of my other instrumental ambient choices for creative energy.
But this album establishes a steady cadence and will keep your foot tapping
while your hands do the typing.