---
title: "Zelda & Chill"
author: "Rich Leyshon"
date: October 8 2023
description: "A Lofi Homage to Nintendo's Zelda Series, Composed by Mikel &
GameChops"
categories:
- Electronic
- Lofi
- Video Game Music
image: https://i.scdn.co/image/ab67616d0000b273fbd9b41a77b7bb918a820541
image-alt: Album cover art
toc: true
---

## Introduction

Nintendo's Zelda series likely needs no introduction. In its fourth decade and
one of the most successful adventure series in video games, Zelda is beloved
for its sense of innovation. 

The music of the games has often been at the core of establishing its identity
and; at times; has played a key construct in the games' development. The N64
batch of games adopted a core game mechanic where the ability to accurately
play songs on an instrument called an ocarina triggered key plot events such as
travelling through time. Perhaps due to this, many of the tracks here are from
this period of Zelda's history.

## Songs for the Recovering Adventurers

Zelda & Chill is a love note to the iconic music produced by the Zelda
developers. As a casual fan, I have enjoyed many of the games in the series and
the album manages to prod the nostalgia sensor in my brain even after multiple
playthroughs. But for those disinterested in video games, this album is  great
for drowning out the outside world and establishing the focus needed for deep
work. 

<iframe style="border-radius:12px" src="https://open.spotify.com/embed/album/3oNO1P0Qlr4oSlMA2MIj67?utm_source=generator&theme=0" width="40%" height="80" frameBorder="0" allowfullscreen="" allow="autoplay; clipboard-write; encrypted-media; fullscreen; picture-in-picture" class=center></iframe>
<br>
Let's take a look at the audio feature analysis for this album. For a formal
description of the audio feature categories presented here, please consult the
<a href="https://developer.spotify.com/documentation/web-api/reference/get-audio-features" target="_blank">Spotify Developer API Documentation</a>.

```{r, width="300px", fig.align="center"}
#| warning: false
#| echo: false
library(here)
library(kableExtra)
library(glue)
source(here("R/music-reviews/spotify-album.R"))
source(here("R/music-reviews/config-secrets.R"))

viz_album_audio_features(
spotify_id = "3oNO1P0Qlr4oSlMA2MIj67",    
title_font_size = 16,
subtitle_font_size = 12,
title_hjust = 0.2,
subtitle_hjust = 0.1,
debug = FALSE
)

```

::: {.callout-tip collapse="true"}
## Interpreting the Ridge Plot (Click to expand)

These ridge plots show a histogram of the mean audio feature values across each
track in the album. Look across the horizontal axis at where the peaks for each
feature occur. For example, if the album contains lots of tracks with busy
vocals, the speechiness feature will show a peak to the right of the horizontal
axis. Conversely, an instrumental album would have a speechiness peak to the
left of the horizontal axis but you would also expect it to have a peak to the
right of the axis in the instrumentalness feature.
:::

I find it surprising that there is such a wide distribution in the valence of
the songs. I expected the album to be distributed centrally as the pace and
tone of the music is generally understated. But there are a couple of tracks
that strike a more melancholy tone, such as Dark World and Gerudo Valley. At
the other end of the scale, the Legend of Zelda theme is likely the most
positive on offer, helping to explain why a broad distribution of valence is
observed. The album's speechiness and instrumentalness score as you'd expect.
Danceability is notably high, potentially as a result of the crunchy drum &
bass lines present in all the songs. Acousticness is generally low throughout
the tracks though with a broad distribution that is potentially influenced by
recurring use of piano and string sections. Perhaps the most interesting
feature for this album is energy, which describes the album as distinctly
middle-of-the-road in terms of intensity. As a chill album you'd likely expect
a lower energy but the strong rhythm section likely lifts the distribution of
this feature.


## The Tracks

Below are the mean audio feature values for each track. The last row (green)
presents a mean summary of the album.

```{r}
#| echo: false
tab = create_album_audio_summary_table(spotify_id = "3oNO1P0Qlr4oSlMA2MIj67")
# create album average
numeric_cols = select(tab, where(is.numeric))
summary_tab = summarise(numeric_cols, across(everything(), mean))
summ_row = cbind(track_name = glue::glue("Album Mean"), summary_tab)
tab = rbind(tab, summ_row)
# Visualise
summ_ind = nrow(tab)
kableExtra::kbl(tab) %>%
row_spec(summ_ind, color = "white", background = "#146542")

```

**Fairy Fountain**

64-bit nostalgia ahead. While the Fairy Fountain has offered sanctuary to any
daring adventurer donning the guise of Link throughout the entire Zelda series,
this theme is synonymous with the N64's Ocarina of Time. This rendition is a
wonderful, crunchy production over a stripped-back ragtime tempo. Warm feelings
of safe havens are immediately conjured from the outset, with that iconic
dreamy harp intro. A strong start to the album that sets the mood for what's to
come.

**Dark world**

For those Zelda fans who can identify with A Link to the Past, this track may
hold more significance. I have not played that game and checking out the
original track, I’d say this version seems like a marked improvement. There's a
fascinating contrast in this song - the meaty bass forms the perfect foil to
the high pitch tubular bell-effect hook.

**Lost Woods**

Another one for the Ocarina of Time fans. A slower tempo than theoriginal,
hence the "Relax" in the album title. There’s a curious reverb effect at play
that sounds a bit like a harpsichord fed through a tin can. After listening to
quite a bit of lofi I am on board with this, but I could see that it may be an
acquired taste for some. Against the light and airy melody, the low-tech
production works really well. 

**Song of Storms**

Familiar to those wishing to summon rain in Ocarina of Time, this laid back
rendition is mixed with a loop of the cooking theme from Breath of the Wild,
replete with periodic pot whistles & cutlery clanking against china. The
creative flair on show as the looped Song of Storms theme builds to a modest
crescendo (it’s chill after all) makes this one of my album highlights. 

**Minuet of Forest**

This song was originally used in Ocarina of Time to warp characters back to the
start of the lost woods play area - one of the more challenging areas in the
early game. The artists have expertly built on what was a very brief movement
in the original material, adding a bouncy low-end bass while keeping the track
structured with a snappy drum section. 

**Gerudo Valley**

Another Ocarina of Time offering here. The original Gerudo Valley theme was
played on Spanish guitar in a flamenco-style with a high tempo. This rendition
 is nearly unrecognisable and undoubtedly brilliant. How the artist has
 isolated the key movements and set them against the slow, chuffing rhythm is
 quite amazing. The song is one of the more melancholy tunes and while lower in
 mood is undoubtedly beautiful.

**Oath to Order**

This song originally appeared in Majora’s Mask and was used as a plot construct
in that game - apparently to summon giants. I have little association with
Majora’s Mask and having listened to it, I fail to identify the musical
connection with the source material. Someone with a greater familiarity would
potentially find more to identify with. As a standalone piece, Oath to Order is
a very warm-sounding, fuzzy sensation with a warbly theremin effect permeating
throughout for a unique twist.

**Dragon Roost Island**

It's great to have a theme from the Windwaker on the album, which in my opinion
is one of the best Zelda games and is often overlooked for its cute exterior.
The original track was distinctly latin with Spanish guitar, maracas and a
forlorn panpipe high in the mix.  As in Gerudo Valley, the artists have slowed
the pace while emphasising the main movement in the original source material.
The guitars have been replaced with warm, fuzzy piano for a more relaxed
atmosphere. 

**Kakariko Village**

Kakariko Village has made recurring appearances throughout the Zelda series of
games and I have struggled to identify which version of the series soundtracks
this song takes its inspiration from. Though the melody is very distinct and
the track is another of my album highlights. I presume it must be from one of
the games I am less familiar with. A light and plucky lick against a bright and
bouncy bassline, there’s plenty to appreciate here.

**Ballad of the Goddess**

Another recurring song in the later games within the Zelda series, I find it a
very enjoyable rendition despite my lack of familiarity with the source
material. The song seems to originate from Skyward Sword, building upon the
melody using sweet harp strings and later a dramatic orchestral accompaniment.
In this version, the artist has retained the harp but has worked with piano for
the main melody. 

**Breath of the Wild**

Taking its inspiration from the main theme of Breath of the Wild which perhaps
lacks a distinctive identity in my opinion, this song starts with a promising
loop of one of its more recognisable movements. However, mid-way through the
song it introduces some distinctly discordant keys that introduce a sort of
dilapidated fairground motif to the song. This is one that I tend to skip for
those reasons - I find it hard to concentrate on my work with discordant sound
in my ears. I view the track as an experiment and not in keeping with the rest
of the album. 

**Hateno Village**

Another Breath of the Wild arrangement, Hateno Village is notably home to the
Ancient Tech Lab - an important site for acquiring gear upgrades as the game
progresses. In this rendition of the theme, the artists have retained some deep
bass strings although the main melody is now expertly played on piano. Again, I
feel that the original material is not the strongest explored in this album,
while the reimagining on offer here is perhaps more interesting than the source
material.

**Legend of Zelda**

Maybe the most iconic video theme of all time. This mix pays homage to the
series’ 8-bit roots by experimenting with a chiptune style that does a great
job of not getting skipped while I work - chiptune can be fun but not generally
for concentration. One of the more moving songs on offer, this track builds to
a crescendo replete with choral and piano arrangements, leaving the listener
with a distinct sense of grandeur capturing the scale of the video games. 

**Ocarina of Time**

The theme from the original N64 Zelda title is a fitting end to this fantastic
homage to perhaps one of the greatest video game adventure series. Kids of my
generation will recall this iconic track as it accompanied the Ocarina of Time
intro. The intro that blew the minds of a generation of Zelda fans who got to
see Link in 3D and on horseback in an amazing open world like never before.
Great times.

## Highs and Lows

The Breath of the Wild track is not to my taste and on writing this review,
I've realised that I consider the soundtrack for this game to be somewhat
inferior to the older games. No hate to Breath of the Wild, which is a
masterpiece in modern gaming, but perhaps the soundtrack is less of a focal
point and potentially appealing to a broader and more mature audience than the
older games in the series.

Standout tracks to my mind are Song of Storms, Kakariko Village and Legend of
Zelda.

## Overall

This album is a great choice for steadying the ship following a busy schedule.
It establishes a quiet focus, allowing your brain's operating system to reboot
following a demanding presentation. I'd suggest this album as the first in a
playlist dedicated to carving out an afternoon of focus time. Set your status
to 'do not disturb' and eliminate your to-do list with the help of Link and his
friends.

For more music to encourage your work efforts, check out
[Productivity Pulse](index.qmd).
