---
title: "Zelda & Chill"
author: "Rich Leyshon"
date: last-modified
description: "A Lofi Homage to Nintendo's Zelda Series, composed by Mikel &
GameChops"
categories:
- Electronic
- Lofi
- Video Game Music
image: https://i.scdn.co/image/ab67616d0000b273fbd9b41a77b7bb918a820541
image-alt: Album cover art
toc: true
---

## Introduction

...

::: {.callout-note}
...
:::

...

## ...

...

<iframe style="border-radius:12px" src="https://open.spotify.com/embed/album/3oNO1P0Qlr4oSlMA2MIj67?utm_source=generator&theme=0" width="40%" height="80" frameBorder="0" allowfullscreen="" allow="autoplay; clipboard-write; encrypted-media; fullscreen; picture-in-picture" class=center></iframe>
<br>
Let's take a look at the audio feature analysis for this album. For a formal
description of the audio feature categories presented here, please consult the
<a href="https://developer.spotify.com/documentation/web-api/reference/get-audio-features" target="_blank">Spotify Developer API Documentation</a>.

```{r, width="300px", fig.align="center"}
#| warning: false
#| echo: false
library(here)
library(kableExtra)
library(glue)
source(here("R/music-reviews/spotify-album.R"))
source(here("R/music-reviews/config-secrets.R"))

viz_album_audio_features(
spotify_id = "3oNO1P0Qlr4oSlMA2MIj67",    
title_font_size = 16,
subtitle_font_size = 12,
title_hjust = 0.2,
subtitle_hjust = 0.1,
debug = FALSE
)

```

::: {.callout-tip collapse="true"}
## Interpreting the Ridge Plot (Click to expand)

These ridge plots show a histogram of the mean audio feature values across each
track in the album. Look across the horizontal axis at where the peaks for each
feature occur. For example, if the album contains lots of tracks with busy
vocals, the speechiness feature will show a peak to the right of the horizontal
axis. Conversely, an instrumental album would have a speechiness peak to the
left of the horizontal axis but you would also expect it to have a peak to the
right of the axis in the instrumentalness feature.
:::

...

## The Tracks

Below are the mean audio feature values for each track. The last row (green)
presents a mean summary of the album.

```{r}
#| echo: false
tab = create_album_audio_summary_table(spotify_id = "3oNO1P0Qlr4oSlMA2MIj67")
# create album average
numeric_cols = select(tab, where(is.numeric))
summary_tab = summarise(numeric_cols, across(everything(), mean))
summ_row = cbind(track_name = glue::glue("Album Mean"), summary_tab)
tab = rbind(tab, summ_row)
# Visualise
summ_ind = nrow(tab)
kableExtra::kbl(tab) %>%
row_spec(summ_ind, color = "white", background = "#146542")

```

**Air**
...

**Infiltrator**
...

**Blood and steel**
...

**Dharma**
...



## Highs and Lows

...
## Overall

...
