---
title: "Triads"
author: "Rich Leyshon"
date: April 19 2024
description: "Code Elektro's exploration of future Japan's criminal underworld"
categories:
- Instrumental
- Synthwave
- Electronic
- Electro
image: https://i.scdn.co/image/ab67616d0000b273c6a43b3a48e548e3da50eeaa
image-alt: Album cover art
toc: true
---

## Dysto-Tokyo

Welcome to neon-lit streets, where the glint of laser-edged katana underlie the
bustling cityscape. Where proud tradition is falling foul of organised
exploitation. A realm where cybernetic dreams and dystopian nightmares converge
in a symphony of pulsating synths and electrifying beats. Code Elektro's
masterpiece third album invites listeners on an immersive journey through a
world where tradition meets technology, and the line between reality and
virtuality blurs. With each track, the album paints a vivid portrait of a
futuristic Tokyo, where shadowy figures roam the alleys and threat stalks the
unwary.


<iframe style="border-radius:12px" src="https://open.spotify.com/embed/album/3K48lplYXbQSc87pfuJ0Pa?utm_source=generator
" width="40%" height="80" frameBorder="0" allowfullscreen="" allow="autoplay; clipboard-write; encrypted-media; fullscreen; picture-in-picture" class=center></iframe>

<br>

For a formal description of the audio feature categories presented here, please
consult the
<a href="https://developer.spotify.com/documentation/web-api/reference/get-audio-features" target="_blank">Spotify Developer API Documentation</a>.

Triads is perfect for those moments when you crave an escape into a futuristic
realm of imagination. Whether you're embarking on a late-night drive through
the city, diving into a cyberpunk novel, or simply seeking to enhance your
creative flow, this album serves as the ideal soundtrack to accompany your
journey into the unknown. Its pulsating rhythms and evocative melodies evoke a
sense of adventure and intrigue, making it the perfect companion for any
escapade into the realm of dystopian dreams.

```{r, width="300px", fig.align="center"}
#| warning: false
#| echo: false
library(here)
library(kableExtra)
library(glue)
source(here("R/music-reviews/spotify-album.R"))
source(here("R/music-reviews/config-secrets.R"))
album_id = "3K48lplYXbQSc87pfuJ0Pa"
viz_album_audio_features(
spotify_id = album_id,    
title_font_size = 14,
subtitle_font_size = 10,
title_hjust = 0.2,
subtitle_hjust = 0.1,
debug = FALSE
)

```

::: {.callout-tip collapse="true"}
## Interpreting the Ridge Plot (Click to expand)

These ridge plots show a histogram of the mean audio feature values across each
track in the album. Look across the horizontal axis at where the peaks for each
feature occur. For example, if the album contains lots of tracks with busy
vocals, the speechiness feature will show a peak to the right of the horizontal
axis. Conversely, an instrumental album would have a speechiness peak to the
left of the horizontal axis but you would also expect it to have a peak to the
right of the axis in the instrumentalness feature.
:::

## The Tracks

Below are the mean audio feature values for each track. The last row (green)
presents a mean summary of the album.

```{r}
#| echo: false
tab = create_album_audio_summary_table(spotify_id = album_id)
# create album average
numeric_cols = select(tab, where(is.numeric))
summary_tab = summarise(numeric_cols, across(everything(), mean))
summ_row = cbind(track_name = glue::glue("Album Mean"), summary_tab)
tab = rbind(tab, summ_row)
# Visualise
summ_ind = nrow(tab)
kableExtra::kbl(tab) %>%
row_spec(summ_ind, color = "white", background = "#146542")

```

The ridge plot profile is remarkably similar to the
[ghostrunner ost](/./music-reviews/01-ghostrunner-ost.qmd), reviewed back in
September 2023. Like ghostrunner, this albunm is low in speechiness and high in
instrumentalness. In contrast, ghostrunner reported a broader distribution in
valence than triads, which I find surprising. Triads is certainly a less
menacing listen with some really beautiful melodies influenced by Japanese
culture.

## Rise of the Triads (Intro)
This introductory track sets the stage for adventure. With pulsating synths and
an aura of anticipation, it serves as the perfect prelude to the sonic journey
ahead.

## Shinobi
"Shinobi" channels the spirit of classic ninja films with its driving rhythms
and mysterious melodies. Its electrifying energy evokes intrigue and images of
stealthy warriors navigating through moonlit landscapes, ready to face any
challenge.

## Night Train
My favourite track. "Night Train" is an inspired prelude to action. An
inspiring ride through the skies of a neon-lit cyberpunk city. A steady tempo
and glittering percussive synth arrangement introduce the city's grand scale
accompanied by an alluring mix of danger and opportunity.

## Chinese Dreams
With "Chinese Dreams," Code Elektro delivers a mesmerizing blend of
Eastern-inspired melodies and retro-futuristic synths. Its hypnotic rhythms
and evocative instrumentation transport listeners to a realm where ancient
traditions meet cutting-edge technology. 

## The Monk
"The Monk" immerses listeners in a meditative soundscape, the onset of a surly
tempest and a gong that could have been sampled straight from a '70s martial
arts flick. Contemplative melodies set to swelling rhythms create a sense of
growing urgency. 

## In the Shadows
"In the Shadows" conjures images of clandestine meetings and covert operations
with its brooding atmosphere and enigmatic melodies. Its pulsating rhythms and
suspenseful build-ups create a sense of tension and intrigue, keeping listeners
on the edge of their seats. Something dark is afoot, and only a select few
will ever know of what happened this day.

## The Wilderness
"The Wilderness" transports listeners to a landscapes where nature has been
thoroughly hybridised with technology. Its expansive soundscapes and ethereal
melodies evoke feelings of awe and wonder, then interspersed with an abundance
of chiptune glitch and reverb, challenging assumptions and provoking wonder.

## International Karate
This track is a steady movement with inspired homages to the sweet melodies of
the East. Its minimalist approach provides a moment of introspection amidst the
album's high-energy tracks.

## Triads
The title track of the album is a tour de force of synth-driven bliss. "Triads"
starts with a subtle ascent evoking impending action or imminent discovery.
Then somewhere a door is kicked through and violence spills out onto the
street, with a bassy reverb drawing the attention of all passers-by to the
apparent chaos. Its pulsating rhythms and euphoric melodies leave a lasting
impression.

## Mission Control
"Mission Control" takes listeners on a cosmic voyage through space and time.
Its propulsive rhythms evoke the spirit of space exploration and the promise of
discovery, inviting listeners to join in on an epic adventure to the far
reaches of the universe.

## Silent Runner
"Silent Runner" captivates listeners with its hauntingly beautiful melodies and
hypnotic rhythms. A sweet soprano accompaniment counters a mysteriously
brooding rhythm section. Its ethereal atmosphere and dreamlike quality create a
sense of weightlessness, like drifting through the cosmos on a silent journey.

## Tokyo Dawn
Closing out the album is "Tokyo Dawn," a mesmerizing journey through to the
dawn, bringing a sense of security and achievement. A sweet and contemplative
outtro accompanying the hero through their jounrey home.

## Overall

In conclusion, "Triads" is a captivating odyssey through a cybernetic
dreamscape, where each track offers a unique sonic experience that transports
listeners to distant realms of imagination. Whether you're a fan of synthwave
or simply appreciate immersive music that takes you on a journey, "Triads" is
an album that is sure to leave a lasting impression.

For more music to encourage your work efforts, check out
[Productivity Pulse](index.qmd).
